Unshuffling Permutations

### 2015-09-12 article.tex
notations

\documentclass[a4paper,10pt]{llncs}

@@

\input{complexity}

@@

We follow the usual terminology on words~\cite{ChoffrutKarhumaki1997}.
Let us recall here the most important ones. The {\em empty word} is
denoted by $\epsilon$. If $u$ is a word, we denote by $\widetilde{u}$
the {\em mirror image} of $u$. If $P$ is a subset of $[|u|]$, $u_{|P}$
is the subword of $u$ consisting in the letters of $u$ at the positions
specified by the elements of $P$. The {\em shuffle} of two words $u$ and
$v$ is the set recursively defined by
$u \shuffle \epsilon = u = \epsilon \shuffle u$ and
$ua \shuffle vb = (u \shuffle vb)a \cup (ua \shuffle v)b$, were $a$ and
$b$ are letters. A word $u$ is a {\em square} if there exists a word $v$
such that $u$ belongs to $v \shuffle v$.
\medskip

We denote by $S_n$ the set of permutations of size $n$ and by $S$ the
set of all permutations. In this paper, permutations are specified by
words without repetition of length $n$ on the alphabet $[n]$, so that
all above definitions about words remain valid on permutations. If
$\pi$ is a permutation, we denote by $\bar \pi$ (resp.~$\pi^{-1}$) the
{\em complementary} (resp. {\em inverse}) of $\pi$. If $u$ is a word of
integers without multiple occurrences of a same letter, $\STD(u)$ is the
{\em standarized} of $u$, that is the unique permutation of the same
size as $u$ such that for all $i, j \in [|u|]$, $u_i < u_j$ if and only
if $\STD(u)_i < \STD(u)_j$. In particular, the image of the map $\STD$
is the set $S$ of all permutations. Two words $u$ and $v$ having the
same standarized are {\em order isomorphic}. If $\sigma$ is a permutation,
there is an {\em occurrence} of (the {\em pattern}) $\sigma$ in $\pi$ if
there is a set $P$ of indexes of letters of $\pi$ such that $\sigma$ and
$\pi_{|P}$ are order isomorphic. When $\pi$ does not admit any occurrence
of $\sigma$, $\pi$ {\em avoids} $\sigma$.
\medskip

>>>>

\documentclass[a4paper]{llncs}

@@

\input{complexity}
\input{graph}

@@

We follow the usual terminology on words~\cite{ChoffrutKarhumaki1997}.
Let us recall here the most important ones. The {\em empty word} is
denoted by $\epsilon$. If $u$ is a word, we denote by $\widetilde{u}$
the {\em mirror image} of $u$. If $P$ is a subset of $[|u|]$, $u_{|P}$
is the subword of $u$ consisting in the letters of $u$ at the positions
specified by the elements of $P$. The {\em shuffle} of two words $u$ and
$v$ is the set recursively defined by
$u \shuffle \epsilon = u = \epsilon \shuffle u$ and
$ua \shuffle vb = (u \shuffle vb)a \cup (ua \shuffle v)b$, were $a$ and
$b$ are letters. A word $u$ is a {\em square} if there exists a word $v$
such that $u$ belongs to $v \shuffle v$.

We denote by $S_n$ the set of permutations of size $n$ and by $S$ the
set of all permutations. In this paper, permutations are specified by
words without repetition of length $n$ on the alphabet $[n]$, so that
all above definitions about words remain valid on permutations. If
$\pi$ is a permutation, we denote by $\bar \pi$ (resp.~$\pi^{-1}$) the
{\em complementary} (resp. {\em inverse}) of $\pi$. If $u$ is a word of
integers without multiple occurrences of a same letter, $\STD(u)$ is the
{\em standarized} of $u$, that is the unique permutation of the same
size as $u$ such that for all $i, j \in [|u|]$, $u_i < u_j$ if and only
if $\STD(u)_i < \STD(u)_j$. In particular, the image of the map $\STD$
is the set $S$ of all permutations. Two words $u$ and $v$ having the
same standarized are {\em order isomorphic}. If $\sigma$ is a permutation,
there is an {\em occurrence} of (the {\em pattern}) $\sigma$ in $\pi$ if
there is a set $P$ of indexes of letters of $\pi$ such that $\sigma$ and
$\pi_{|P}$ are order isomorphic. When $\pi$ does not admit any occurrence
of $\sigma$, $\pi$ {\em avoids} $\sigma$.

For a graph $G$, we denote $\VS{G}$ as the set of vertices and
$\ES{G}$ as the set of edges.
Let $u = u_1 \, u_2 \,\ldots\, u_n$ be a word.
The \emph{clique graph associated to $u$}, denoted $K_u$,
is the complete graph on $\VS{K_u} = \{u_1, u_2, \ldots, u_n\}$.
The structure of this underlying graph is linear,
\emph{i.e.}, the set of vertices is equipped with a natural total order
$<$ defined by $u_i < u_j$ if and only if $i < j$.
(For a linear graph, we write $(u_i, u_j)$ instead of $\{u_i, u_j\}$
for an edge if it is clear from the context that $i < j$.)
Recall that two edges of a graph are \emph{independent} if they do not share
a common vertex, and that a \emph{matching} $\mathcal{M}$ in $G$ is a set of
pairwise independent edges.
A matching is \emph{perfect} if it covers all the vertices of the graph.
In case the set of vertices is equipped with a total order,
a matching $\mathcal{M}$ is said to be \emph{containment-free}
if there do not exist (independent) edges $(u_i, u_j)$ and
$(u_k, u_\ell)$ in $\mathcal{M}$ such that
$u_i < u_k < u_\ell < u_j$ or  $u_k < u_i < u_j < u_\ell$.
